---
title: "Lab 0"
author: "Mu Lu"
date: "2015-09-21"
output: html_document
---

## The Dataframe

Read in your data set and assign to the object:

```{r}
present <- read.table("http://assets.datacamp.com/course/dasi/present.txt")
```

`present` will print your data set in the console:

```{r, echo=FALSE}
present
```

Print the number of rows and variables:
```{r}
dim(present)
```

Print the names of the variables of the data frame:
```{r}
names(present)
```

We can tell there are 3 vaibales (year,boys,girls) in the present data set.

Find the number of boys and girls born each year, and assign answers to variables:
```{r}
num_boys <-present$boys
num_girls <- present$girls
num_boys
num_girls
```

The result is in data type **vector**.

## Arithmetic calculation on dataframe and plottting

Plot the data:

```{r}
plot(x=present$year,y=present$girls,xlab = "Year" ,ylab = "Girls Born")
```

Line plot:

```{r}
plot(x=present$year,y=present$girls,xlab = "Year" ,ylab = "Girls Born", type = "l")
```

Plotting the total number of new born babies by year:
```{r}
babies <- present$boys + present$girls
plot(x=present$year,y=present$girls,xlab = "Year" ,ylab = "Babies Born", type = "l")
```

Compute the ratio of the number of boys divided by the number of girls born:
```{r}
present$boys/present$girls
```

Calculate the proportion of male newborns for all years and plot it:
```{r}
plot(x=present$year,y=present$boys/(present$boys + present$girls),xlab = "Year" ,ylab = "Boys Born", type = "l")
```

In addition to simple mathematical operators like subtraction and division, you can ask R to make comparisons: greater than with `>`, less than with `<`, and equality with `==`. We can ask if boys outnumber girls in each year:
```{r}
present$boys > present$girls
```

Plotting boys-to-girls ration for every year:
```{r}
plot(x=present$year,y=present$boys/present$girls,xlab = "Year" ,ylab = "Boys vs girls ration", type = "l")
```

Plotting absolute differences between number of boys and girls born in each year:
```{r}
plot(x=present$year,y=abs(present$boys-present$girls),xlab = "Year" ,ylab = "Boys/girls diff", type = "l")
```